---
title: "Development actions history"
output: html_document
editor_options: 
  chunk_output_type: console
---

All commands that you use to use when developing packages...

# First time just after creating the project

- Fill the following chunk to create the DESCRIPTION of your package

```{r description, eval=FALSE}
# Describe your package
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "Create Reproducible Sensory Reports",
    Description = "A Shiny app to query a database in which QDA sensory data are stored, analyse data R package and generate reproducible and interactive HTML sensory reports. Such tool can enable sensory analysts and panel leaders to spend less time preparing reports, and add more value to their work.",
    `Authors@R` = c(
      person("Margot", "Brard", email = "margot@thinkr.fr", role = c("aut", "cre"), comment = c(ORCID = "0000-0001-6754-0659")),
      person(given = "ThinkR", role = "cph")
    )
  )
)
# Define License with use_*_license()
usethis::use_mit_license("Margot Brard")
```

# Start using git

```{r, eval=FALSE}
usethis::use_git()
# Deal with classical files to ignore
usethis::git_vaccinate()
# Use main for primary branch
usethis::git_default_branch_rename()
```

# Set extra sources of documentation

```{r, eval=FALSE}
# Install a first time
remotes::install_local()
# README
usethis::use_readme_rmd()
# Code of Conduct
usethis::use_code_of_conduct("contact@fake.com")
# NEWS
usethis::use_news_md()
```

**From now, you will need to "inflate" your package at least once to be able to use the following commands. Let's go to your flat template, and come back here later if/when needed.**


# Package development tools
## Use once

```{r, eval=FALSE}
# Pipe
usethis::use_pipe()

# package-level documentation
usethis::use_package_doc()

# Add new flat template
fusen::add_flat_template(
  flat_name = "database", 
  template = "additional"
)
```

## Use everytime needed

```{r}
# Simulate package installation
pkgload::load_all()

# Generate documentation and deal with dependencies
attachment::att_amend_desc()

# Check the package
devtools::check()
```

# Share the package

```{r}
# set and try pkgdown documentation website
usethis::use_pkgdown()
pkgdown::build_site()

# build the tar.gz with vignettes to share with others
devtools::build(vignettes = TRUE)
```

